---
format: html
---

```{r}
library(tidyverse)
library(nycflights13)
```

## Exercise / Examples 

1. How many flights to Los Angeles (LAX) did each of the legacy carriers (AA, UA, DL or US) have in May from JFK, and what was their average duration?

```{r}
flights |>
  filter(dest == "LAX") |> 
  filter(carrier %in% c("AA","UA","DL","US")) |>
  filter(month == 5) |>
  filter(origin == "JFK") |>
  summarize(
    n = n(),
    avg_duration = mean(air_time, na.rm=TRUE),
    .by = carrier
  )
```

2. What was the shortest flight out of each airport in terms of distance? In terms of duration?

```{r}
flights |>
  group_by(origin) |>
  filter(air_time == min(air_time, na.rm=TRUE)) |>
  select(origin, dest, air_time)
```


```{r}
flights |>
  group_by(origin) |>
  slice_min(air_time, n=5) |>
  select(origin, dest, air_time)
```

```{r}
flights |>
  group_by(origin) |>
  slice_min(distance, n=1) |>
  select(origin, dest, distance) |>
  distinct()
```

3. Which plane (check the tail number) flew out of each New York airport the most?

```{r}
flights |>
  count(origin, tailnum) |>
  filter(!is.na(tailnum)) |>
  slice_max(n, n=1, by=origin)
```


4. Which date should you fly on if you want to have the lowest possible average departure delay? What about arrival delay?

```{r}
flights |>
  summarize(
    avg_dep_delay = mean(dep_delay, na.rm = TRUE),
    .by = c(month, day, origin)
  ) |>
  slice_min(avg_dep_delay, n = 1, by = origin)
```

```{r}
flights |>
  summarize(
    avg_arr_delay = mean(arr_delay, na.rm = TRUE),
    .by = c(month, day, origin)
  ) |>
  slice_min(avg_arr_delay, n = 1, by = origin)
```
